Kaggle-in-class Data Challenges Can Boost Student Learning

### 2020-08-11 paper-kaggle.Rmd

Figure \ref{fig:Monash_quiz} displays the quiz scores taken by the undergraduate (ST-UG) and postgraduate (ST-PG) students, as a beeswarm plot [@beeswarm]. The distribution of quiz scores are similar, except that undergaduates tend to have a longer lower tail.

@@

Students were randomly assigned to one of two groups, regression (R) or classification (C). Each group will have a data competition designed for that topic, made available through kaggle in class (https://inclass.kaggle.com). Students wored individually to build predictive models, and submit predictions for the first 16 days. On the 16^{th} day of the competition students received an email that they allowed to establish groups and continue the competition as a group. The competition was continued for additional 7 days. In case they chose to form a group, the final score of the group in total was accounted as their individual mark towards the final grade for the subject and analysed in this project.  Only 3 groups were formed, two groups with 3 students and one group with two students. All other students chose to continue the competition individually.

@@

Figure \ref{fig:crossperformances} presents students' scores for classification and regression questions. A score over 1 is considered outperformance (relative to the expectation). Quarters one and three include students that underperform or outperform on both types of questions, respectively. In both cases the number of students that participated in the classification competition is very close to the number of students that participated in the regression competition (excluding a few regression students on the border of score 1). Students in quarters two and four outperform on one type of questions but not on the other type. We can see that more regression students outperform on regression questions than classification students 
(`r filter(t, Proj_alloc == "regression" ,     outperf_class == FALSE,  outperf_reg == TRUE)["n"]`<!-- 12 --> vs. 
 `r filter(t, Proj_alloc == "classification" , outperf_class == FALSE,  outperf_reg == TRUE)["n"]`<!-- 7 -->). 
Similarly, classification students do better on classification questions 
(`r filter(t, Proj_alloc == "classification" , outperf_class == TRUE,  outperf_reg == FALSE)["n"]`<!-- 11 --> vs. 
 `r filter(t, Proj_alloc == "regression" ,     outperf_class == TRUE,  outperf_reg == FALSE)["n"]`<!-- 3 -->). 
This is another evidence towards positive influence of the data competition on student's performances.

@@

It may be recommended to limit students to one submission per day. It encourages students to think about more efficient improvement of their model before the next submission. It also prevents the student spending too much time building and submitting models. Some students will become so engaged in the competition that they might neglect their other coursework. About halfway through the coompetition, students might be allowed to form teams, to learn how averaging models can boost performance.

@@

This paper has described the setup and results of an experiment to examine the effectiveness of data competitions on student learning, using Kaggle InClass as the vehicle for conducting the competition. The experiment was conducted in the classroom settings as part of the normal teaching of the courses, which imposes limitations on the design. However, with both CSDM and ST, either a randomized assignment of students to two topic groups, or a control group, was possible, which enabled comparing performance.

Running an experiment in the classroom setting imposes, of course, some limitations. One of them is a relatively small sample size of each group. The other is the need to address the university policies, such as the group assessment approach we took for CSDM.

One of the major limitations is the ability to include such a control group that it is easy to show that its members are equal to the case group members. In the case of Statistical Thinking subject the control group was taken as the undergraduate students that took this subject. At first, using undergraduate students as a control group for graduate students may be surprising.  However, the experience of teaching this subject over several years and some statistical comparison of the two groups can justify this decision. Undergraduate students that took this subject are strong students. Their performance in other tasks and exam questions, not relevant to the competition, was equivalent to the postgraduate students cohort.

The primary finding, is that participating in a data challenge competition,  produces a statistically significant improvement in the learning of the topic, although the effect size is small.  Secondarily, the competitions enhanced interest and engagement in the course.

>>>>

Figure \ref{fig:Monash_quiz} displays the quiz scores taken by the undergraduate (ST-UG) and postgraduate (ST-PG) students, as a beeswarm plot [@beeswarm]. The distribution of quiz scores are similar, except that undergraduates tend to have a longer lower tail.

@@

Students were randomly assigned to one of two groups, regression (R) or classification (C). Each group will have a data competition designed for that topic, made available through kaggle in class (https://inclass.kaggle.com). Students worked individually to build predictive models, and submit predictions for the first 16 days. On the 16^{th} day of the competition students received an email that they allowed to establish groups and continue the competition as a group. The competition was continued for additional 7 days. In case they chose to form a group, the final score of the group in total was accounted as their individual mark towards the final grade for the subject and analysed in this project.  Only 3 groups were formed, two groups with 3 students and one group with two students. All other students chose to continue the competition individually.

@@

Figure \ref{fig:crossperformances} presents students' scores for classification and regression questions. A score over 1 is considered as outperform (relative to the expectation). Quarters one and three include students that underperform or outperform on both types of questions, respectively. In both cases the number of students that participated in the classification competition is very close to the number of students that participated in the regression competition (excluding a few regression students on the border of score 1). Students in quarters two and four outperform on one type of questions but not on the other type. We can see that more regression students outperform on regression questions than classification students 
(`r filter(t, Proj_alloc == "regression" ,     outperf_class == FALSE,  outperf_reg == TRUE)["n"]`<!-- 12 --> vs. 
 `r filter(t, Proj_alloc == "classification" , outperf_class == FALSE,  outperf_reg == TRUE)["n"]`<!-- 7 -->). 
Similarly, classification students do better on classification questions 
(`r filter(t, Proj_alloc == "classification" , outperf_class == TRUE,  outperf_reg == FALSE)["n"]`<!-- 11 --> vs. 
 `r filter(t, Proj_alloc == "regression" ,     outperf_class == TRUE,  outperf_reg == FALSE)["n"]`<!-- 3 -->). 
This is another evidence towards positive influence of the data competition on student's performances.

@@

It may be recommended to limit students to one submission per day. It encourages students to think about more efficient improvement of their model before the next submission. It also prevents the student spending too much time building and submitting models. Some students will become so engaged in the competition that they might neglect their other coursework. About halfway through the competition, students might be allowed to form teams, to learn how averaging models can boost performance.

@@

This paper has described an experiment to examine the effectiveness of data competitions on student learning, using Kaggle InClass as the vehicle for conducting the competition. The primary finding, is that participating in a data challenge competition,  produces a statistically significant improvement in the learning of the topic, although the effect size is small.  Secondarily, the competitions enhanced interest and engagement in the course.

Because the experiment was conducted in the classroom setting as part of the normal teaching of the courses, which imposed limitations on the design. However, performance comparison was enabled in CSDM by a randomized assignment of students to two topic groups, and in ST by using a control group. In CDSM, the group sizes were relatively small, approximately 30 students per group. For ST the control group was the undergraduate students that took class. Using undergraduate students as a control group for graduate students may be surprising.  However, the experience of teaching this subject over several years and some statistical comparison of the two groups justifies the approach. Undergraduate students performance in other tasks and exam questions, not relevant to the competition, was equivalent to the postgraduate students cohort.
